---
title: "Create a palette from an image using clustering"
format: html
editor: visual
---

The idea here is to use K-means clustering to identify centroids in RGB space based on an image.  By selecting different numbers of clusters, you can create palettes of different size.

This was (very roughly) inspired by the [wesanderson](https://github.com/karthik/wesanderson) and [calecopal](https://github.com/an-bui/calecopal) packages, and also cool image manipulation examples like https://www.r-bloggers.com/2022/10/probabilistic-photograph-manipulation-with-ggplot2-and-imager. 

Once the base code is up and running, perhaps turn it into a Shiny app where users can load their own image and set up their own palettes.

```{r setup}
library(imager)
library(tidyverse)
```

The first step is to load in an image, and convert it to a dataframe of `x`, `y`, `r`, `g`, `b`.

Note: for a Shiny app, might be nice to have an option to downsample a really large image.

```{r}
img <- load.image("20221112_DSC_0068_TP.JPG") # load image in
dims <- dim(img)[1:2] # get dimensions for exporting
# change to data frame
img_dat <- img %>% 
  as.data.frame(wide = "c") %>% 
  mutate(hex = rgb(c.1, c.2, c.3), xy = paste(x, y, sep = "_"))

```
